
```{r, include = FALSE}
library(CLARiTI)
library(arsenal)
library(ggplot2)
library(tidyverse)
library(nlme)

data(investigator_scan_mrisbm_nacc66)
data(investigator_scan_taupetnpdka_nacc66)
data(investigator_scan_amyloidpetgaain_nacc66)
data(NACC_ADSP_PHC_Cognition_2024)
data(investigator_ftldlbd_nacc66)

knitr::opts_chunk$set(echo = TRUE,
  collapse = TRUE,
  comment = "#>",
  fig.width = 8, 
  fig.height = 8 / 1.6, 
  out.width = "100%",
  cache = FALSE
)

options(digits = 2, width = 150)

theme_set(theme_minimal())

scale_colour_discrete <-
    function(...) ggsci::scale_color_lancet()
scale_fill_discrete <-
    function(...) ggsci::scale_fill_lancet()
```

# Organize data

Then for cognition, you could use the harmonized cognitive scores from Tim's group.  Those are here. 

However, it's possible there is a lot of missingness in Tim's harmonized scores.  If so, then you could use individual tests from the larger UDS file.

1.	PHC_MEM 		Harmonized Composite Memory Score 
2.	PHC_EXF		Harmonized Composite Executive Function Score 
3.	PHC_LAN		Harmonized Composite Language Score

```{r}
# For hippocampus volume from SCAN
dd <- investigator_scan_mrisbm_nacc66 %>%
  select(NACCID, DATE = SCANDT, HIPPOCAMPUS, ICV = CEREBRUMTCV) %>%
  full_join(investigator_scan_taupetnpdka_nacc66 %>%
    select(NACCID, DATE = SCANDATE, Tau_PET = META_TEMPORAL_SUVR, TRACER) %>%
    mutate(TRACER = case_when(
      TRACER == 1 ~ 'FDG',
      TRACER == 2 ~ 'PIB',
      TRACER == 3 ~ 'Florbetapir',
      TRACER == 4 ~ 'Florbetaben',
      TRACER == 5 ~ 'NAV4694',
      TRACER == 6 ~ 'Flortaucipir',
      TRACER == 7 ~ 'MK6240',
      TRACER == 8 ~ 'PI2620',
      TRACER == 9 ~ 'GTP1',
      TRACER == 99 ~ 'Unknown') %>% as.factor()) %>%
    rename(Tau_TRACER = TRACER),
    by = c('NACCID', 'DATE')) %>%
  full_join(investigator_scan_amyloidpetgaain_nacc66 %>%
    select(NACCID, DATE = SCANDATE, AMYLOID_STATUS, CENTILOIDS, TRACER) %>%
    mutate(TRACER = case_when(
      TRACER == 1 ~ 'FDG',
      TRACER == 2 ~ 'PIB',
      TRACER == 3 ~ 'Florbetapir',
      TRACER == 4 ~ 'Florbetaben',
      TRACER == 5 ~ 'NAV4694',
      TRACER == 6 ~ 'Flortaucipir',
      TRACER == 7 ~ 'MK6240',
      TRACER == 8 ~ 'PI2620',
      TRACER == 9 ~ 'GTP1',
      TRACER == 99 ~ 'Unknown') %>% as.factor()) %>%
    rename(Amyloid_TRACER = TRACER),
    by = c('NACCID', 'DATE')) %>%
  full_join(NACC_ADSP_PHC_Cognition_2024 %>%
    select(NACCID, NACCVNUM, PHC_MEM, PHC_EXF, PHC_LAN) %>%
    left_join(investigator_ftldlbd_nacc66 %>%
      mutate(DATE = as.Date(paste(VISITYR, VISITMO, VISITDAY, sep='-'))) %>%
      select(NACCID, NACCVNUM, DATE),
      by = c('NACCID', 'NACCVNUM')),
    by = c('NACCID', 'DATE')) %>%
  left_join(investigator_ftldlbd_nacc66 %>%
    mutate(BIRTHDATE = as.Date(paste(BIRTHYR, BIRTHMO, 15, sep = '-'))) %>%
    filter(!is.na(BIRTHDATE)) %>%
    arrange(NACCID, NACCVNUM) %>%
    select(NACCID, BIRTHDATE, RACE, SEX, EDUC, HISPANIC) %>%
    group_by(NACCID) %>%
    fill(.direction = "updown") %>%
    ungroup() %>%
    filter(!duplicated(NACCID)),
    by = 'NACCID') %>%
  mutate(Age = as.numeric(DATE - BIRTHDATE)/365.25) %>%
  group_by(NACCID) %>%
  mutate(ICV = mean(ICV, na.rm = TRUE)) %>%
  ungroup() %>%
  arrange(NACCID, DATE)

dd_cross <- dd %>%
  group_by(NACCID) %>%
  fill(everything(), .direction = "updown") %>%
  filter(!duplicated(NACCID)) %>%
  mutate(
    SEX = case_when(is.na(SEX) ~ " Missing", TRUE ~ SEX) %>%
      factor(levels = c(" Male", "Female", " Missing"))
  ) %>%
  ungroup()

# harmonize tau PET data ----
tmp <- dd %>% filter(!is.na(Tau_PET) & !is.na(Age))
tau_PET_fit <- lme(Tau_PET ~ Tau_TRACER + SEX + Age, data = tmp,
  random = ~ 1 | NACCID,
  weights = varIdent(form = ~ 1 | Tau_TRACER))
tmp$Tau_PET_ComBat <- ComBat(Tau_PET ~ Tau_TRACER + SEX + Age, data = tmp,
  random1 = ~ 1 | NACCID, random2 = NULL,
  weights = varIdent(form = ~ 1 | Tau_TRACER))

dd <- dd %>%
  left_join(tmp %>% select(NACCID, DATE, Tau_PET_ComBat), 
    by = c('NACCID', 'DATE'))
```

# Baseline characteristics tables

```{r}
mylabels = list(
  PHC_MEM = 'Harmonized Memory', 
  PHC_EXF = 'Harmonized Exec Function', 
  PHC_LAN = 'Harmonized Language',
  EDUC = 'Education (yrs)',
  Age = 'Age (yrs)')
```

## All CLARiTI participants
```{r results = 'asis', include = TRUE}
tableby(SEX ~ Age + RACE + HISPANIC + 
    AMYLOID_STATUS + CENTILOIDS +
    HIPPOCAMPUS + Tau_PET + Tau_TRACER +
    PHC_MEM + PHC_EXF + PHC_LAN + EDUC, 
  data = dd_cross %>% filter(SEX != ' Missing')) %>%
  summary(labelTranslations = mylabels, digits = 1)
```

## CLARiTI participants with hippocampal volumes
```{r results = 'asis', include = TRUE}
tableby(SEX ~ Age + RACE + HISPANIC + 
    AMYLOID_STATUS + CENTILOIDS +
    HIPPOCAMPUS + Tau_PET + Tau_TRACER +
    PHC_MEM + PHC_EXF + PHC_LAN + EDUC, 
  data = dd_cross %>% filter(!is.na(HIPPOCAMPUS) & SEX != ' Missing')) %>%
  summary(labelTranslations = mylabels, digits = 1)
```

## CLARiTI participants with tau PET
```{r results = 'asis', include = TRUE}
tableby(SEX ~ Age + RACE + HISPANIC + 
    AMYLOID_STATUS + CENTILOIDS +
    HIPPOCAMPUS + Tau_PET + Tau_TRACER +
    PHC_MEM + PHC_EXF + PHC_LAN + EDUC, 
  data = dd_cross %>% filter(!is.na(Tau_PET) & SEX != ' Missing')) %>%
  summary(labelTranslations = mylabels, digits = 1)
```

## CLARiTI participants with amyloid PET
```{r results = 'asis', include = TRUE}
tableby(SEX ~ Age + RACE + HISPANIC + 
    AMYLOID_STATUS + CENTILOIDS +
    HIPPOCAMPUS + Tau_PET + Tau_TRACER +
    PHC_MEM + PHC_EXF + PHC_LAN + EDUC, 
  data = dd_cross %>% filter(!is.na(CENTILOIDS) & SEX != ' Missing')) %>%
  summary(labelTranslations = mylabels, digits = 1)
```

# Summary plots

```{r CLARiTI-spaghetti, fig.cap='Spaghetti of selected outcomes by age and race'}
dd %>% 
  select(NACCID, Age, RACE, HIPPOCAMPUS, Tau_PET_ComBat, CENTILOIDS,
    PHC_MEM, PHC_EXF, PHC_LAN) %>%
  pivot_longer(HIPPOCAMPUS:PHC_LAN) %>%
  filter(!is.na(value) & !is.na(RACE)) %>%
ggplot(aes(x=Age, y=value, color=RACE)) +
  geom_point(alpha=0.1) +
  geom_line(aes(group = NACCID), alpha=0.01) +
  facet_wrap(vars(name), scales = 'free_y') +
  guides(colour = guide_legend(override.aes = list(alpha=1))) +
  ylab('')
```

```{r CLARiTI-loess, fig.cap='Locally weighted scatterplot smooths (LOWESS) of selected outcomes by age and race'}
dd %>% 
  select(NACCID, Age, RACE, HIPPOCAMPUS, Tau_PET_ComBat, CENTILOIDS,
    PHC_MEM, PHC_EXF, PHC_LAN) %>%
  pivot_longer(HIPPOCAMPUS:PHC_LAN) %>%
  filter(!is.na(value) & !is.na(RACE)) %>%
ggplot(aes(x=Age, y=value, color=RACE)) +
  geom_smooth(method = 'lm') +
  # geom_smooth(method = 'gam', formula = y ~ s(x, bs = "cs", fx = TRUE, k = 1)) +
  facet_wrap(vars(name), scales = 'free_y') +
  guides(colour = guide_legend(override.aes = list(alpha=1))) +
  ylab('')
```

# ComBat Harmonization of tau PET (tracers)

```{r}
summary(tau_PET_fit)
ggplot(dd %>% filter(!is.na(Tau_PET_ComBat)), 
  aes(x = Tau_PET, y = Tau_PET_ComBat, color = Tau_TRACER)) +
  geom_point() +
  geom_abline(intercept = 0, slope = 1, linetype = 'dashed')
```
