---
title: "CLARiTI Summaries"
author: "ATRI Biostatistics"
output:   
  rmarkdown::html_vignette:
    toc: true
bibliography: references.bib  
urlcolor: blue
vignette: >
  %\VignetteIndexEntry{CLARiTI Summaries}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

The **Consortium for Clarity in ADRD Research Through Imaging (CLARiTI)** [@mormino2025consortium] leverages existing ADRC infrastructure to integrate standardized imaging and plasma collection to characterize mixed pathologies and use community-engaged research methods to ensure that ≥ 25% of the sample is from underrepresented populations (e.g., ethnoculturally minoritized, low education). 

The code below demonstrates how to use data in the `CLARITI` package and @R to derive some basic summaries.

```{r, include = FALSE}
library(CLARiTI)
library(arsenal)
library(ggplot2)
library(tidyverse)
library(nlme)

data(investigator_scan_mrisbm_nacc66)
data(investigator_scan_taupetnpdka_nacc66)
data(investigator_scan_amyloidpetgaain_nacc66)
data(NACC_ADSP_PHC_Cognition_2024)
data(investigator_ftldlbd_nacc66)

knitr::opts_chunk$set(echo = TRUE,
  collapse = TRUE,
  comment = "#>",
  fig.width = 8, 
  fig.height = 8 / 1.6, 
  out.width = "100%",
  cache = FALSE
)

options(digits = 2, width = 150)

theme_set(theme_minimal())

scale_colour_discrete <-
    function(...) ggsci::scale_color_lancet()
scale_fill_discrete <-
    function(...) ggsci::scale_fill_lancet()
```

```{r, child='../man/rmd/CLARiTI-Summaries-child.Rmd'}
```

# References
